---
title: "Portfolio CM"
author: "Jonna"
date: "9-2-2021"
output:
  flexdashboard::flex_dashboard:
    storyboard: true
---




```{r include=FALSE}
library(tidyverse)
library(spotifyr)
library(compmus)
library(formattable)
library(plotly)
```

```{r include=FALSE}
# Downloading the playlists from Spotify

tchai <- get_playlist_audio_features("", "37i9dQZF1DX2kvBlhyEkb6")
mahler <- get_playlist_audio_features("", "37i9dQZF1DWYFH7pZ5xe7h")

```

```{r include=FALSE}
# Adding the playlists together into one dataframe

music <-
  bind_rows(
    tchai %>% mutate(category = "Tchaikovsky"),
    mahler %>% mutate(category = "Mahler")
  )

df <- data.frame(music)

```


### Showing the *chords* of **Nutcracker** and **

```{r, echo=FALSE}
circshift <- function(v, n) {
  if (n == 0) v else c(tail(v, n), head(v, -n))
}

#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

rev_major_key <-
  c(5.0, 2.0, 3.5, 2.0, 4.5, 4.0, 2.0, 4.5, 2.0, 3.5, 1.5, 4.0)
rev_minor_key <-
  c(5.0, 2.0, 3.5, 4.5, 2.0, 4.0, 2.0, 4.5, 3.5, 2.0, 1.5, 4.0)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )
  
rev_key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(rev_major_key, 6),
    "Bb:min", circshift(rev_minor_key, 10),
    "Db:maj", circshift(rev_major_key, 1),
    "F:min", circshift(rev_minor_key, 5),
    "Ab:maj", circshift(rev_major_key, 8),
    "C:min", circshift(rev_minor_key, 0),
    "Eb:maj", circshift(rev_major_key, 3),
    "G:min", circshift(rev_minor_key, 7),
    "Bb:maj", circshift(rev_major_key, 10),
    "D:min", circshift(rev_minor_key, 2),
    "F:maj", circshift(rev_major_key, 5),
    "A:min", circshift(rev_minor_key, 9),
    "C:maj", circshift(rev_major_key, 0),
    "E:min", circshift(rev_minor_key, 4),
    "G:maj", circshift(rev_major_key, 7),
    "B:min", circshift(rev_minor_key, 11),
    "D:maj", circshift(rev_major_key, 2),
    "F#:min", circshift(rev_minor_key, 6),
    "A:maj", circshift(rev_major_key, 9),
    "C#:min", circshift(rev_minor_key, 1),
    "E:maj", circshift(rev_major_key, 4),
    "G#:min", circshift(rev_minor_key, 8),
    "B:maj", circshift(rev_major_key, 11),
    "D#:min", circshift(rev_minor_key, 3)
  )
```

```{r}

# CHORDOGRAMS

#Symphony no. 5 in C sharp minor: 4. Adagietto (sehr langsam)

#With cosine distance and Euclidean norm


no5_chord <- get_tidy_audio_analysis("0M78i2tXl07RGPjcm0InHA") %>%
  compmus_align(bars, segments) %>%
  select(bars) %>%
  unnest(bars) %>%
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "mean", norm = "manhattan"
      )
  ) 

no5_chord %>% 
  compmus_match_pitch_template(chord_templates, "cosine", "euclidean") %>%
  ggplot(
    aes(x = start + duration / 2, width = duration, y = name, fill = d)
  ) +
  geom_tile() +
  scale_fill_viridis_c(guide = "none") +
  theme_minimal() +
  labs(x = "Time (s)", y = "")




```




```{r}
#Nutcracker Op. 71, Act 2: No. 13. Waltz of the flowers

nuts_chord <- get_tidy_audio_analysis("0CJuLeZpCdqrqoCMXWZw8k") %>%
  compmus_align(bars, segments) %>%
  select(bars) %>%
  unnest(bars) %>%
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "mean", norm = "manhattan"
      )
  ) 

nuts_chord %>% 
  compmus_match_pitch_template(chord_templates, "cosine", "euclidean") %>%
  ggplot(
    aes(x = start + duration / 2, width = duration, y = name, fill = d)
  ) +
  geom_tile() +
  scale_fill_viridis_c(guide = "none") +
  theme_minimal() +
  labs(x = "Time (s)", y = "")



```


```{r}

# KEYGRAM

# Tchaikovsky - Nutcracker

nuts_keys <-
  get_tidy_audio_analysis("0CJuLeZpCdqrqoCMXWZw8k") %>%
  compmus_align(sections, segments) %>%
  select(sections) %>%
  unnest(sections) %>%
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "mean", norm = "manhattan"
      )
  )

#revised

nuts_keys %>% 
  compmus_match_pitch_template(
    rev_key_templates,         # Change to chord_templates if desired
    method = "cosine",  # Try different distance metrics
    norm = "euclidean"     # Try different norms
  ) %>%
  ggplot(
    aes(x = start + duration / 2, width = duration, y = name, fill = d)
  ) +
  geom_tile() +
  scale_fill_viridis_c(guide = "none") +
  theme_minimal() +
  labs(x = "Time (s)", y = "")



# Mahler- No5


no5_keys <-
  get_tidy_audio_analysis("0M78i2tXl07RGPjcm0InHA") %>%
  compmus_align(sections, segments) %>%
  select(sections) %>%
  unnest(sections) %>%
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "mean", norm = "manhattan"
      )
  )


#revised


no5_keys %>% 
  compmus_match_pitch_template(
    rev_key_templates,         # Change to chord_templates if desired
    method = "cosine",  # Try different distance metrics
    norm = "euclidean"     # Try different norms
  ) %>%
  ggplot(
    aes(x = start + duration / 2, width = duration, y = name, fill = d)
  ) +
  geom_tile() +
  scale_fill_viridis_c(guide = "none") +
  theme_minimal() +
  labs(x = "Time (s)", y = "")


```



```{r}


bind_rows(
  nuts_chord %>%
    compmus_match_pitch_template(
  chord_templates, 
  "cosine", 
  "euclidean") %>%
    mutate(type= "Nutcracker chordogram"),
  no5_chord %>%
    compmus_match_pitch_template(
  chord_templates, 
  "cosine", 
  "euclidean") %>%
    mutate(type= "Mahler no. 5 chordogram"),
  nuts_keys %>%
    compmus_match_pitch_template(
    rev_key_templates,         # Change to chord_templates if desired
    method = "cosine",  # Try different distance metrics
    norm = "euclidean"     # Try different norms
  ) %>%
    mutate(type= "Nutcracker keygram"),
  no5_keys %>%
    compmus_match_pitch_template(
    rev_key_templates,         # Change to chord_templates if desired
    method = "cosine",  # Try different distance metrics
    norm = "euclidean"     # Try different norms
  ) %>%
    mutate(type= "Mahler no. 5 keygram"),
) %>%
  mutate() %>%
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = name,
      fill = d
    )
  ) +
  geom_tile() +
  facet_wrap(~type, scales="free")+
  labs(x = "Time (s)", y = "") +
  scale_fill_viridis_c(guide= "none") +                              
  theme_classic() 


```




### Looking at the underlying structures of our favorite tracks with a **ceptogram**

```{r}

# 

swan <-
  get_tidy_audio_analysis("37XdVwSMJQ4EVBrLFcZGyx") %>% # Change URI.
  compmus_align(bars, segments) %>%                     # Change `bars`
  select(bars) %>%                                      #   in all three
  unnest(bars) %>%                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

swanSections <-
  get_tidy_audio_analysis("37XdVwSMJQ4EVBrLFcZGyx") %>% # Change URI.
  compmus_align(sections, segments) %>%                     # Change `bars`
  select(sections) %>%                                      #   in all three
  unnest(sections) %>%                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

no3 <-
  get_tidy_audio_analysis("4oXa32SjjuVrBBjEMm94Mx") %>% # Change URI.
  compmus_align(bars, segments) %>%                     # Change `bars`
  select(bars) %>%                                      #   in all three
  unnest(bars) %>%                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

no3Sections <-
  get_tidy_audio_analysis("4oXa32SjjuVrBBjEMm94Mx") %>% # Change URI.
  compmus_align(sections, segments) %>%                     # Change `bars`
  select(sections) %>%                                      #   in all three
  unnest(sections) %>%                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )
```

```{r}

bind_rows(
  swan %>%
    compmus_gather_timbre() %>%
    mutate(type= "Swan Lake per bar"),
  no3 %>%
    compmus_gather_timbre() %>%
    mutate(type= "Mahler no. 3 per bar"),
  swanSections %>%
    compmus_gather_timbre() %>%
    mutate(type= "Swan Lake per section"),
  no3Sections %>%
    compmus_gather_timbre() %>%
    mutate(type= "Mahler no. 3 per section"),
) %>%
  mutate() %>%
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = basis,
      fill = value
    )
  ) +
  geom_tile() +
  facet_wrap(~type, scales="free")+
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  scale_fill_viridis_c(guide= "none") +                              
  theme_classic()


```

***
These graphs show cepstrograms for *Swan Lake, Op.20, Act 2: No. 10* (the Swan theme) and *Symphony no. 3 in D Minor: V.*. 

It is clear that there is a high magnitude for the 3rd coefficient throughout both tracks. However, with these Mel-frequency cepstral coefficients it is not entirely clear what the third coefficient exactly represents. If one was to listen to the track whilst following along with the cepstogram, it seems as though the magnitude for co3 is the highest at the moments when the theme is played in high tones. 

For *Swan Lake*, this is when the oboe that plays the melody of the track. Once the rest of the orchestra joins in, the magnitude diminishes somewhat. At this point, the music also speeds up. At the end, there is another spike of magnitude when the other instruments fall away and the oboe is again clearly audible. When the oboe dies down, there also is a spike in co2, when the lower toned wind instruments finish the theme. 
For *No. 3*, coefficient 3 is the highest when the woodwinds blow their highest tones during the melody around the 20th second mark and at the end around 230-250 seconds. At these moments, the woodwinds are most clearly audible. It therefore seems that coefficient 3 for both tracks is related to the high tones of these instruments. 

In the 'sections' graphs, it is clear that for *no. 3* these tones are present  throughout the track and especially in the first and last parts. For *Swan Lake*, these high tones are mostly present in the first part of the track, and gradually diminish. 

### The structures of the Swan theme made visible with **chroma** and **timbre**

```{r}

maria <-
  get_tidy_audio_analysis("37XdVwSMJQ4EVBrLFcZGyx") %>%
  compmus_align(bars, segments) %>%
  select(bars) %>%
  unnest(bars) %>%
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "acentre", norm = "manhattan"
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "mean"
      )
  )
bind_rows(
  maria %>%
    compmus_self_similarity(pitches, "aitchison") %>%
    mutate(d = d / max(d), type = "Chroma"),
  maria %>%
    compmus_self_similarity(timbre, "cosine") %>%
    mutate(d = d / max(d), type = "Timbre")
) %>%
  mutate() %>%
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")
```

***
In order to look further into Swan Lake, I have created self-similarity matrices demonstrating its chroma and timbre features.
  
  In the chroma-based self-similarity matrix, there are no parallel diagonal lines visible. This means that there are no exact repetitions in the track. Even though the theme of the track seems to repeats itself, this is done by various different instruments and with slightly different melodies wrought trough it. Therefore, we can see that there are no exact repetitions of the theme in the chromagram. Around the 75-95 into the song, the matrix shows sudden yellow vertical lines. This points to a sudden musical novelty in the song, and in this case it marks the moment when the string orchestra starts playing a melody that is different from the theme of the song.
  
  The timbre-based self-similarity matrix shows a clear divide of the track in 3 main sections. The first section from 0-50 seconds represents the moment when the oboe is almost solely audible. Hereafter, there is a section from 50-75 seconds, which is when the other wind instruments join into the theme. During the final section hereafter, the string orchestra takes over the theme. Finally, the most striking aspect of this matrix is the sudden burst of yellow at the end of the song, around the 150 second mark. This is the moment when the lower toned wind instruments finish the theme, as this is the only moment in the track when we can solely hear these lower tones. In the matrix, we can also see these tones fading out a little bit.

### The subjects of this research: **Tchaikovsky** and **Mahler**

For this assignment I will be conducting an exploratory analysis of the works of Pyotr Ilyich Tchaikovsky (1840-1893) to Gustav Mahler (1860-1911). My reasons for choosing to compare these two artists are mostly personal. I do not know much about classical music, as I mostly tend to listen to contemporary music. An exception to this rule is Tchaikovsky, whose musical pieces I have enjoyed listening to ever since my mother and grandmother took me to see *Swan Lake* when I was young. When I was working on my thesis in the summer of 2020, I started listening to classical music more often and found it quite enjoyable and relaxing. However, I still do not know much about this genre of music and I certainly do not feel like I can identify the composer of a musical piece on the radio just by listening to it, as my mother can. Therefore, I wanted to take this opportunity to dive more deeply into this genre of music in order to increase my knowledge on its different musical elements and the ways in these can be used by composers to create their own distinct styles. Besides Tchaikovksy, I also chose to analyze works by Mahler, as this is my mother's favorite composer. It would be interesting to see in what ways the styles of our favorite composers differ. Furthermore, Tchaikovsky's pieces are seen as works of the Romantic period, whereas Mahler is seen as a mix of the 19th century Austro-German music tradition and the modernism of the early 20th century. I selected the works that I will analyze by adding the Spotify public playlists "This is Tchaikovsky" (60 tracks) and "This is Mahler" (70 tracks) together. I chose these two playlists, as they were created by Spotify in order to represent work typical for the composers. They are performed by a variety of orcestra's, which means that there coudl be some deviations based on the perfomers' interpretations. I will pay special attention to mine and my mother's favorite tracks of these composers. For me, these are *Swan Lake, Op.20, Act 2: No. 10* and *The Nutcracker, Op 71, Act 2: No. 13*, which coincidentally are also the most popular Tchaikovsky tracks. For my mother, these are  *Symphony No. 5 in C-sharp Minor: IV* and *Symphony no. 3 in D Minor: V.*. 

### The overall **differences** between these two composers



```{r echo=FALSE, fig.cap = "Figure 1. Table comparing means and modes of Tchaikovsky and Mahler"}
# Creating a function to find the statistical mode

getmode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}


# Creating a table comparing means and modes

MeanTable <- df %>%
  group_by(Composer = category) %>%
  summarise(Danceability = mean(danceability),
            Energy = mean(energy),
            Key = getmode(key),
            Loudness = mean(loudness),
            Mode = getmode(mode),
            Valence = mean(valence),
            Tempo= mean(tempo)
            )

formattable(MeanTable)

```


```{r include=FALSE}
# Creating a dataframe of means to be used in the plot

GraphMeans <- df %>%
  group_by(category) %>%
  summarise(valence = mean(valence),
            energy = mean(energy))


```




```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.cap= "Figure 2. Interactive plot comparing Tchaikovsky and Mahler"}
# Creating a plot comparing valence, energy and danceability of Tchaikovsky and Mahler 

plot1 <-music %>% 
  ggplot(aes(x = valence, y = energy, color=category, size=danceability)) + 
  geom_point(shape=17, alpha=0.6, aes(text = paste(
               "Composer: ", category, "\n",
               "Track: ", track.name, "\n",
               "Valence: ", valence, "\n",
               "Energy: ", energy, "\n",
               "Danceability: ", danceability, "\n"
            ))) +
  geom_point(data=GraphMeans, size=6, aes(text= paste(
    "Composer: ", category, "\n",
    "Mean Valence: ", format(round(valence, 2), nsmall=2), "\n",
    "Mean Energy: ", format(round(energy, 2), nsmall=2), "\n"
  ))) +
  geom_smooth(se=F) +
  theme_classic() +
  theme(legend.title=element_blank())
  labs(x = "Valence",
       y = "Energy",
       title = "Interactive plot comparing Tchaikovsky and Mahler",
       size = "Danceability")

  
ggplotly(plot1, tooltip="text")

```

***
I first created a table of the means of the variables danceability, energy, loudness, valence and tempo of both composers, to which I also added the statistical mode's for key and mode. Tchaikovsky seems to be higher overall in danceability, energy, valence and tempo, whereas Mahler's pieces have been played more loudly. Furthermore, Tchaikovsky has most often composed pieces in Key=2, meaning D (also C-double sharp, E-double flat) . For Mahler, the key in which he most often composed is F (also E-sharp, G-double flat). Both composers have most often composed in major.

In the graph underneath, I plotted the tracks for Tchaikovsky and Mahler on their valence and energy. The size of the triangles indicates the measure of danceability, with larger triangles meaning more danceable tracks. The trend for both composers is shown in the lines in the graph. The line representing Tchaikovsky lies higher than the line for Mahler, and the former also reaches further with valence. However, this far reach of the line is based mostly on one point, *Dance des petits cygnes* in *Swan Lake*. and is therefore not necessarily representable for his whole body of work. I also plotted the means for both composers' valence and energy on the graph. Here we can see more clearly that both have composed pieces that are generally low in valence and energy. Looking at the distribution of points, it seems that Mahler especially has the greatest portion of his work in this area, with Tchaikovsky having a somewhat larger amount of variation. Furthermore, there seems to be quite a large variation in danceability, as the triangles vary in sizes. The most danceable tracks also have the highest valence. However there doesn't seem to be a further correlation between valence and danceability from looking at this graph alone, as there are also tracks with higher than average danceability for lower amounts of valence.


